---
title: "STT 461 Project"
author: "Dylan Mather, Caitlyn Locke, Colin Richard"
date: "4/20/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,include = FALSE)
```

```{r}
library(tidyverse)
```


```{r}


pst_data <- read.csv("pst_data.csv")
pst_data$date <- as.Date(with(pst_data, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

mt_data <- read.csv("mt_data.csv")
mt_data$date <- as.Date(with(mt_data, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

est_data <- read.csv("est_data.csv")
est_data$date <- as.Date(with(est_data, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")

ct_data <- read.csv("ct_data.csv")
ct_data$date <- as.Date(with(ct_data, paste(Year, Month, Day,sep="-")), "%Y-%m-%d")


```

```{r}
pst_data <-pst_data%>%
  arrange(date)
head(pst_data)
tail(pst_data)

ts_rest <- ts(pst_data$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult <- decompose(ts_rest, type = "multiplicative")


plot(ts_rest)

plot(decomp_mult$trend)
plot(decomp_mult$seasonal)


```
```{r,include = FALSE}
pst_data%>%
  filter(Year == c(2000:2002))%>%
  filter(AvgTemperature > -50)%>%
  ggplot(aes(date,AvgTemperature)) + geom_point()+geom_smooth()
pst_data%>%
  filter(Year == c(2000:2004))%>%
  filter(AvgTemperature > -50)%>%
  ggplot(aes(date,AvgTemperature)) + geom_point()+geom_smooth()
pst_data%>%
  filter(Year == c(2000:2006))%>%
  filter(AvgTemperature > -50)%>%
  ggplot(aes(date,AvgTemperature)) + geom_point()+geom_smooth()
```






Pacific Data
```{r}
pst_avg_temp <- pst_data%>%
  filter(AvgTemperature > -50)%>%
  group_by(date)%>%
  summarise(r_avg_temp = mean(AvgTemperature),year = mean(Year),day = mean(Day),month = mean(Month))
  


ts_rest_PST <- ts(pst_avg_temp$r_avg_temp, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_PST <- decompose(ts_rest_PST, type = "multiplicative")



plot(decomp_mult_PST$seasonal)


```


```{r}
train <- pst_avg_temp%>%
  filter(year < 2008)

test <- pst_avg_temp%>%
  filter(year >=2008)



ccf(train$date,train$r_avg_temp)
ccf(train$day,train$r_avg_temp)
ccf(train$month,train$r_avg_temp)
ccf(train$year,train$r_avg_temp)

fit_pst = lm(r_avg_temp ~ date +lag(month,300) + day ,data = train)
summary(fit_pst)


pred_fit <- predict(fit_pst,test)

plot(test$date,test$r_avg_temp)
lines(test$date,pred_fit,col = "red",lwd=2)

plot(pred_fit,test$r_avg_temp)

```






Hypothesis Testing for PST
```{r}
avg_tmp_pst_96_10 <- pst_avg_temp%>%
  filter(year > 1996 & year < 2010)%>%
  select(r_avg_temp)


avg_tmp_pst_12_15 <- pst_avg_temp%>%
  filter(year > 2012 & year < 2015)%>%
  select(r_avg_temp)


t.test(avg_tmp_pst_12_15,avg_tmp_pst_96_10, alternative = "greater")
```




Eastern Data

```{r}
est_avg_temp <- est_data%>%
  filter(AvgTemperature > -50)%>%
  group_by(date)%>%
  summarise(r_avg_temp = mean(AvgTemperature),year = mean(Year),day = mean(Day),month = mean(Month))


ts_rest_EST <- ts(est_avg_temp$r_avg_temp, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_EST <- decompose(ts_rest_EST, type = "multiplicative")


```

Hypothesis Testing for EST
```{r}
avg_tmp_est_96_10 <- est_avg_temp%>%
  filter(year > 1996 & year < 2010)%>%
  select(r_avg_temp)


avg_tmp_est_12_15 <- est_avg_temp%>%
  filter(year > 2012 & year < 2015)%>%
  select(r_avg_temp)


t.test(avg_tmp_est_12_15,avg_tmp_est_96_10, alternative = "greater")
```


Central Data

```{r}
ct_avg_temp <- ct_data%>%
  filter(AvgTemperature > -50)%>%
  group_by(date)%>%
  summarise(r_avg_temp = mean(AvgTemperature),year = mean(Year),day = mean(Day),month = mean(Month))


ts_rest_CT <- ts(ct_avg_temp$r_avg_temp, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_CT <- decompose(ts_rest_CT, type = "multiplicative")

```

Hypothesis Testing for CT
```{r}
avg_tmp_ct_96_10 <- ct_avg_temp%>%
  filter(year > 1996 & year < 2010)%>%
  select(r_avg_temp)


avg_tmp_ct_12_15 <- ct_avg_temp%>%
  filter(year > 2012 & year < 2015)%>%
  select(r_avg_temp)


t.test(avg_tmp_ct_12_15,avg_tmp_ct_96_10, alternative = "greater")
```


Mountain data


```{r}
mt_avg_temp <- mt_data%>%
  filter(AvgTemperature > -50)%>%
  group_by(date)%>%
  summarise(r_avg_temp = mean(AvgTemperature),year = mean(Year),day = mean(Day),month = mean(Month))

ts_rest_MT <- ts(mt_avg_temp$r_avg_temp, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_MT <- decompose(ts_rest_MT, type = "multiplicative")

```

Hypothesis Testing for MT
```{r}
avg_tmp_mt_96_10 <- mt_avg_temp%>%
  filter(year > 1996 & year < 2010)%>%
  select(r_avg_temp)


avg_tmp_mt_12_15 <- mt_avg_temp%>%
  filter(year > 2012 & year < 2015)%>%
  select(r_avg_temp)


t.test(avg_tmp_mt_12_15,avg_tmp_mt_96_10, alternative = "greater")
```


Fresno Data
```{r}
Fresno_temp <- pst_data%>%
  filter(City == "Fresno" & AvgTemperature > -50)

ts_rest_fresno <- ts(Fresno_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_fesno <- decompose(ts_rest_fresno, type = "multiplicative")

```



Los Angeles Data
```{r}
LA_temp <- pst_data%>%
  filter(City == "Los Angeles" & AvgTemperature > -50)

ts_rest_LA <- ts(LA_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_LA <- decompose(ts_rest_LA, type = "multiplicative")

```


Sacramento
```{r}
Sac_temp <- pst_data%>%
  filter(City == "Sacramento" & AvgTemperature > -50)

ts_rest_Sac <- ts(Sac_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Sac <- decompose(ts_rest_Sac, type = "multiplicative")

```



San Diego Data
```{r}
SD_temp <- pst_data%>%
  filter(City == "San Diego" & AvgTemperature > -50)

ts_rest_SD <- ts(SD_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_SD <- decompose(ts_rest_SD, type = "multiplicative")

```


San Francisco Data
```{r}
SF_temp <- pst_data%>%
  filter(City == "San Francisco" & AvgTemperature > -50)

ts_rest_SF <- ts(SF_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_SF <- decompose(ts_rest_SF, type = "multiplicative")

```



Reno Data
```{r}
Reno_temp <- pst_data%>%
  filter(City == "Reno" & AvgTemperature > -50)

ts_rest_Reno <- ts(Reno_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Reno <- decompose(ts_rest_Reno, type = "multiplicative")

```


Las Vegas Data
```{r}
LV_temp <- pst_data%>%
  filter(City == "Las Vegas" & AvgTemperature > -50)

ts_rest_LV <- ts(LV_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_LV <- decompose(ts_rest_LV, type = "multiplicative")


```

Eugene Data
```{r}
Eug_temp <- pst_data%>%
  filter(City == "Eugene" & AvgTemperature > -50)

ts_rest_Eug <- ts(Eug_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Eug <- decompose(ts_rest_Eug, type = "multiplicative")

```

Medford Data
```{r}
Med_temp <- pst_data%>%
  filter(City == "Medford" & AvgTemperature > -50)

ts_rest_Med <- ts(Med_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Med <- decompose(ts_rest_Med, type = "multiplicative")


```


Portland Data
```{r}
Port_temp <- pst_data%>%
  filter(City == "Portland" & AvgTemperature > -50)

ts_rest_Port <- ts(Port_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Port <- decompose(ts_rest_Port, type = "multiplicative")

```



Salem Data
```{r}
Sal_temp <- pst_data%>%
  filter(City == "Salem" & AvgTemperature > -50)

ts_rest_Sal <- ts(Sal_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Sal <- decompose(ts_rest_Sal, type = "multiplicative")


```

Seattle Data
```{r}
Sea_temp <- pst_data%>%
  filter(City == "Seattle" & AvgTemperature > -50)

ts_rest_Sea <- ts(Sea_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Sea <- decompose(ts_rest_Sea, type = "multiplicative")


```


Spokane Data
```{r}
Spok_temp <- pst_data%>%
  filter(City == "Spokane" & AvgTemperature > -50)

ts_rest_Spok <- ts(Spok_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Spok <- decompose(ts_rest_Spok, type = "multiplicative")

```

Yakima Data
```{r}
Yaki_temp <- pst_data%>%
  filter(City == "Yakima" & AvgTemperature > -50)


ts_rest_Yaki <- ts(Yaki_temp$AvgTemperature, start = c(1995,01,01), end =c(2019,12,31), frequency = 365 )

decomp_mult_Yaki <- decompose(ts_rest_Yaki, type = "multiplicative")


```


```{r}
ts.plot(decomp_mult_PST$trend, decomp_mult_EST$trend, decomp_mult_CT$trend,
        decomp_mult_MT$trend, col = c("red","green","blue","purple"), 
        main = "Temperature Trends for the Time Zones in USA",
        ylab = "Temperature (F)", lwd = c(3,1,1,1))
legend("topl", legend=c("PST", "EST","MT","CT"), ncol = 4,
       col=c("red", "green","purple","blue"), lty=1, cex=0.8,lwd = 2)
```


```{r}
ts.plot(decomp_mult_SD$trend,decomp_mult_LA$trend,decomp_mult_fesno$trend,
        decomp_mult_Sac$trend, decomp_mult_SF$trend, col = c("red","green","blue",
                                                       "purple", "orange"),
        main = "Temperature Trends for California, USA",
        ylab = "Temperature (F)")
legend("topl", legend=c("SD", "LA", "Fresno", "Sacramento",
                        "SF"), ncol = 3,
       col=c("red","green","blue","purple","orange"), lty=1, cex=0.8, lwd = 2)
```
```{r}
ts.plot(decomp_mult_Eug$trend, decomp_mult_Med$trend,
        decomp_mult_Port$trend, decomp_mult_Sal$trend, col = c("red","green","purple","blue"),
        main = "Temperature Trends for Oregon, USA",
        ylab = "Temperature (F)")
legend("topl", legend=c("Eugene", "Medford", "Portland", "Salem"), ncol = 2,
       col=c("red","green","purple","blue"), lty=1, cex=0.8, lwd = 2)
```


```{r}
ts.plot(decomp_mult_Sea$trend,decomp_mult_Spok$trend,
        decomp_mult_Yaki$trend, col = c("red","green","blue"),
        main = "Temperature Trends for Washington, USA",
        ylab = "Temperature (F)")
legend("topl", legend=c("Seattle","Spokane","Yakima"), ncol = 4,
       col=c("red","green","blue"), lty=1, cex=0.8, lwd = 2)
```

```{r}
ts.plot(decomp_mult_LV$trend, decomp_mult_Reno$trend, col = c("red","green"),
        main = "Temperature Trends for Nevada, USA",
        ylab = "Temperature (F)")
legend("topl", legend=c("Las Vegas", "Reno"), ncol = 2,
       col=c("red","green","purple"), lty=1, cex=0.8, lwd = 2)
```

The most recent El Ninos were in 2002–03, 2004–05, 2006–07, 2009–10, 2014–16, and 2018–19
The one starting in 2014 was the longest being 2 years long which most likely accounts for the large spike in the
Pacific data around the same time.


